---
title: "Depth-dependent genetic structuring of *Montastraea cavernosa* and their Symbiodiniaceae communities varies between Alacranes and Bajos del Norte reefs, Campeche Bank, Mexico"
author: "Alexis Sturm"
date: "8/31/2021"
output: 
html_document:
    theme: flatly
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.align = 'left')
#knitr::opts_knit$set(root.dir = '../revisedData')
options(width = 88, scipen = 4)
```

***

# About this document
***  
All analyses were performed in R version 3.6.2.
This is the code that accompanies the publication *Depth-dependent genetic structuring of Montastraea cavernosa and their Symbiodiniaceae communities varies between Alacranes and Bajos del Norte reefs, Campeche Bank, Mexico*. Here you will find all the code to repeat the statistical analyses performed in R and the figures created for the manuscript. The accompanying library preparation protocol and bioinformatic walkthrough can be found by clicking on the links.

If you have any issues or questions about the code please feel free to send an email to lexie.sturm@gmail.com.

***

# Basic setup of R environment
***
## Loading required packages
Make sure to set the working directory:
`setwd("~/path/to/directory/with/data")`

```{r, set working directory}
setwd('/Users/student/Documents/GitHub/mxMcavSnp/')
```

For the following analyses we will require the use of multiple different R packages, we can use the package *pacman* to quickly load them.
```{r, load packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("adegenet", "dendextend", "flextable", "gdata", "ggdendro", "hierfstat", "Imap", "kableExtra", 
               "paletteer", "patchwork", "officer", "poppr", "RColorBrewer", "reshape2", "StAMPP", 
               "tidyverse", "vcfR", "vegan", "WGCNA", "ggnewscale", "rnaturalearth", "sf", "ggspatial", 
               "cowplot", "marmap", "dplyr", "edgeR", "MCMC.OTU", "pairwiseAdonis", "distGeo")

pacman::p_load_gh("eliocamp/ggnewscale", "ropensci/rnaturalearthhires")
```

# Figures and analyses
##Sampling Map
```{r, sampling map}
mxSites = read.csv("mxSampleSites.csv", header = TRUE)
mxSites$depthZone=as.factor(mxSites$depthZone)
# coerce to sf object

mxSites1 <- st_as_sf(x = mxSites, coords = c('lon', 'lat'), crs = 4326)
mxSitesDist=st_distance(mxSites1)
mxSitesDistKm=mxSitesDist/1000
rownames(mxSitesDistKm) <-c(mxSites$reef)
colnames(mxSitesDistKm) <-c(mxSites$reef)
#mxSites$depth = factor(mxSites$depth, levels = levels(mxSites$depth)[c(2,1)])
#mxSites$reef = factor(mxSites$reef, levels = levels(mxSites$reef)[c(1, 7, 8, 3, 6, 2, 5, 4, 11, 10, 9, 12)])
mxReefs = mxSites %>% group_by(reef) %>% summarize(lat = mean(lat), lon = mean(lon))
#mxReefs$depthShape=c("Mesophotic", "Shallow", "Shallow", "Shallow", "Shallow", "Shallow", "Shallow", "Shallow", "Paired", "Paired", "Paired", "Paired")
#mxReefs$depthShape=as.factor(mxReefs$depthShape)
#mxReefs$depthShape = factor(mxReefs$depthShape, levels = levels(mxReefs$depthShape)[c(2,3,1)])
gomCountries = st_as_sf(ne_countries(country = c("United States of America", "Cuba", "Mexico", "The Bahamas", "Belize", "Guatemala", "Cayman Islands"), scale = "Large"))


gom5.100Bathy=st_read("/Users/student/Documents/GitHub/mxMcavSnp/gom_5-100m/w98e78n31s18_isobath_5-100m.shp") %>% st_transform(54030)
gom200Bathy=st_read("/Users/student/Documents/GitHub/mxMcavSnp/GOM_isobath_100-1000m/w98e78n31s18_isobath_100-1000m.shp") %>%
filter(CONTOUR == "-200") %>% st_transform(54030)

mxPal=c("seagreen3", "blue")

mxMapA = ggplot() +
  geom_sf(data = gomCountries, fill = "white", size = 0.25) +
  geom_sf(data = gom200Bathy, color = "red") +
  geom_point(data = mxSites, aes(x = lon, y = lat, shape = depthZone), size = 0) +
  geom_point(data = mxReefs, aes(x = lon, y = lat, fill = reef), size = 3, shape=21) +
  scale_fill_manual(values = mxPal, name = "Site") +
  scale_shape_manual(values = c(24,22,23,25), breaks=c("10","15", "25", "35"), name = "Depth Zone", labels=c("10 m","15 m", "25 m", "35 m")) +
  #scale_color_manual(values = c("blue"), labels = c("MPA Boundaries")) + #name = "Boundaries",
  coord_sf(xlim = c(-103, -80), ylim = c(18.5, 30)) +
  scale_x_continuous(breaks = c(seq(-103, -80, by = 2))) +
  scale_y_continuous(breaks = c(seq(18.5, 30, by = 2))) +
  #geom_rect(aes(xmin = -89.8, xmax = -89.6, ymin = 22.35, ymax = 22.6), color = "black", fill = NA, size = 0.4) +
  #geom_rect(aes(xmin = -88.74, xmax = -88.68, ymin = 23.24, ymax = 23.32), color = "black", fill = NA, size = 0.4) +
  annotation_scale(location = "bl") +
  annotation_north_arrow(location = "tr", which_north = "true", style = north_arrow_minimal()) +
  guides(fill = guide_legend(override.aes = list(shape = 22, color = NA, size = 4), ncol = 2, order = 1), shape = guide_legend(override.aes = list(size = 3), order = 2)) +
  #guides(fill = guide_legend(override.aes = list(shape = 21, size = 4), order = 1))+ #,
         #shape = guide_legend(order=2),
         #color = guide_legend(title=NULL, order = 3))+ #override.aes = list(fill=NA),
  theme_bw() +
  theme(plot.title = element_text(size = 9) ,
        panel.background = element_rect(fill = "aliceblue"),
        plot.background = element_blank(),
        axis.title = element_blank(),
        legend.position = "bottom")

#mxMapA

#ggsave("mxMapNoInset.tiff", plot = mxMapA, width = 30, height = 20, units = "cm", dpi = 600)
set.seed(224)
inset = ggplot() +
  geom_segment(aes(x = -103, xend = -80, y = 18.5, yend = 30), color = "gray92", size = .55) +
  geom_sf(data = gomCountries, fill = "white", size = 0.25) +
  geom_point(data = subset(mxSites, subset = mxSites$reef == "Alacranes"), aes(x = lon, y = lat, fill = reef, shape=depthZone), size = 3, position=position_jitter(width=0.01, height=0.01)) +
  geom_point(data = subset(mxSites, subset = mxSites$reef == "Bajos del Norte"), aes(x = lon, y = lat, fill = reef, shape=depthZone), size = 3, position=position_jitter(width=0.002, height=0.002)) +
  scale_fill_manual(values = mxPal, name = "Reef") +
  scale_shape_manual(values = c(24,22,23,25), breaks=c("10","15", "25", "35"), name = "Depth Zone", labels=c("10 m","15 m", "25 m", "35 m")) +
  #geom_jitter()+
  #geom_sf(data = gom5.100Bathy, color = "gray") +
  guides(fill = guide_legend(override.aes = list(shape = 21, color = NA, size = 4))) +
  theme_bw() +
  theme(legend.title = element_text(size = 9, face = "bold"),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "aliceblue"),
        legend.text = element_text(size = 9),
        legend.position = "none",
        legend.direction = "vertical",
        legend.box = "vertical",
        panel.grid=element_blank(),
        rect = element_blank())

# inset

alacranes = inset+
  ggtitle("Alacranes") +
  # geom_sf(data = gomCountries, fill = "white", size = 0.25) +
  annotation_scale(location = "bl") +
  coord_sf(xlim = c(-89.8, -89.6), ylim = c(22.35, 22.6)) +
  scale_x_continuous(breaks = c(seq(-89.8, -89.6, by = .1))) +
  scale_y_continuous(breaks = c(seq(22.35, 22.6, by = .1)))+
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.background=element_blank(),
        panel.border = element_rect(size = 1))

# alacranes

bajosDelNorte = inset +
  ggtitle("Bajos del Norte") +
  annotation_scale(location = "bl") +
  coord_sf(xlim = c(-88.74, -88.68), ylim = c(23.24, 23.32)) +
  scale_x_continuous(breaks = c(seq(-88.74, -88.68, by = .1))) +
  scale_y_continuous(breaks = c(seq(23.24, 23.32, by = .1)))+
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.background=element_blank(),
        plot.title = element_text(vjust=-6.5, hjust=0.5),
        panel.border = element_rect(size = 1))

#bajosDelNorte

mxMap = ggdraw() +
  draw_plot(mxMapA) +
  #draw_plot(inset) +
  draw_plot(alacranes, x = 0.01, y = 0.25, width = 0.3, height = 0.3) +
  draw_plot(bajosDelNorte, x = 0.67, y = 0.25, width = 0.3, height = 0.3)


mxMap

ggsave("./figures/fig1.tiff", plot = mxMap, width = 25, height = 20, units = "cm", dpi = 600)
```

##Host PCoA with IBS
```{r, Host PCoA with IBS}
snpMa = as.matrix(read.table("mxMcavNoClones.ibsMat"))
mxMds = cmdscale(snpMa, eig = TRUE, x.ret = TRUE)

# Determine percent variation captured on each axis
# Calculate the eigenvalues so later we can figure out % variation shown on each Principal Coordinate
mxSnpPcoaVar = round(mxMds$eig/sum(mxMds$eig)*100, 1)
mxSnpPcoaVar

# Format data to plot
mxSnpPcoaValues = mxMds$points
mxSnpPcoaValues

snpI2P = read.csv("mxInds2PopsNoClones.csv") # 2-column tab-delimited table of individual assignments to populations; must be in the same order as samples in the bam list or vcf file.
row.names(snpI2P) = snpI2P[,1]
mxSnpPcoaValues=cbind(snpI2P, mxSnpPcoaValues)
mxSnpPcoaValues =as.data.frame(mxSnpPcoaValues, sample = rownames(mxSnpPcoaValues))
colnames(mxSnpPcoaValues)[5] <- "PCo1"
colnames(mxSnpPcoaValues)[6] <- "PCo2"
mxSnpPcoaValues

snpPCoA = merge(mxSnpPcoaValues, aggregate(cbind(mean.x=PCo1,mean.y=PCo2)~popDepth, mxSnpPcoaValues, mean), by="popDepth")

snpPCoA$depthZoneM=as.factor(snpPCoA$depthZoneM)
#snpPCoA$depthZone = factor(snpPCoA$depthZone, levels(snpPCoA$depthZone)[c(2,1)])

# SNP PCoA biplot
mxSnpPcoaPlotA = ggplot(snpPCoA, aes(x = PCo1, y = PCo2, color = site, fill = site, shape = depthZoneM, linetype = depthZoneM)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_vline(xintercept = 0, color = "gray90", size = 0.5) +
  #stat_ellipse(data = subset(snpPCoA, type = "n", geom = "polygon")) + #ellipse
  #scale_linetype_manual(values=c(1,2,4,3), name = "Depth Zone")+
  geom_point(aes(x = PCo1, y = PCo2, shape = depthZoneM), size = 3, alpha = 0.3, show.legend = FALSE, guide=FALSE) + #individual's points indicated by circles
  scale_shape_manual(values = c(24,22,23,25), breaks=c("10","15", "25", "35"), name = "Depth Zone", labels=c("10 m","15 m", "25 m", "35 m")) +
  geom_point(aes(x = mean.x, y = mean.y, shape = depthZoneM), size = 5, color = "black") + #population centroids indicated by triangles
  #scale_fill_paletteer_d("LaCroixColoR::PassionFruit", breaks=c("Alacranes", "Bajos del Norte"), name= "Population")+
  scale_fill_manual(values = c("seagreen3", "blue"), name= "Population")+
  #scale_color_paletteer_d("LaCroixColoR::PassionFruit", breaks=c("Alacranes", "Bajos del Norte"), name= "Population", guide=FALSE) +
  scale_color_manual(values = c("seagreen3", "blue"), name= "Population")+
  xlab(paste ("PCo 1 (", mxSnpPcoaVar[1],"%)", sep = "")) + #Prints percent variation explained by first axis
  ylab(paste ("PCo 2 (", mxSnpPcoaVar[2],"%)", sep = "")) + #Prints percent variation explained by second axis
  guides(shape = guide_legend(order = 2), fill = guide_legend(override.aes = list(shape = 22, size = 4, color = NA), order = 1))+ #linetype = guide_legend(order = 3),
  theme_bw()

mxSnpPcoaPlot = mxSnpPcoaPlotA +
  theme(axis.title.x = element_text(color = "black", size = 10),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.title.y = element_text(color = "black", size = 10),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        legend.position = "left",
        panel.border = element_rect(color = "black", size = 1.2),
        panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())



mxSnpPcoaPlot

ggsave("./figures/fig2.tiff", plot = mxSnpPcoaPlot, height = 5, width = 7, units = "in", dpi = 300)
```

## AMOVA and Pairwise Fst
```{r, AMOVA and Pairwise Fst, fig.width = 13.4, fig.height = 5.9}
mxVcf = read.vcfR("mxMcavNoClones.bcf")
mxGenlightPopDepth = vcfR2genlight(mxVcf, n.cores = 2) # Converts the vcf file into a file format that poppr uses the "genlight" format
locNames(mxGenlightPopDepth) = paste(mxVcf@fix[,1],mxVcf@fix[,2],sep="_")
popData = read.csv("mxInds2PopsNoClones.csv") # Reads in population data for each sample

strata(mxGenlightPopDepth) = data.frame(popData)
setPop(mxGenlightPopDepth) = ~popDepth
amova <- poppr.amova(mxGenlightPopDepth, ~popDepth) #Runs AMOVA
amova
set.seed(1999)
amovasignif <- randtest(amova, nrepet = 99) #Calculates significance levels of the AMOVA with 99 permutations
amovasignif
plot(amovasignif)

#mxGenlightPopDepth$pop=as.factor(mxGenlightPopDepth$pop)
#mxGenlightPopDepth$pop = factor(mxGenlightPopDepth$pop, levels(mxGenlightPopDepth$pop)[c(4,3,2,1,5,6,8,7)])

set.seed(694)
mx.fst <- stamppFst(mxGenlightPopDepth, nboots = 99, percent = 95, nclusters = 4) #99 permutations
mx.fst$Fsts
mx.fst$Pvalues

# pca.1 <- glPca(mxGenlightPopDepth, nf=300, n.cores=1) 
# 
# # proportion of explained variance by first three axes
# pca.1$eig[1]/sum(pca.1$eig) # proportion of variation explained by 1st axis
# pca.1$eig[2]/sum(pca.1$eig) # proportion of variation explained by 2nd axis 
# pca.1$eig[3]/sum(pca.1$eig) # proportion of variation explained by 3rd axis
# 
# #####K-Means Clustering and DAPC
# grp <- find.clusters(mxGenlightPopDepth, max.n.clust=11, glPca = pca.1, perc.pca = 100, n.iter=1e6, n.start=1000) 
# 

pop.order <- c("Alacranes-10", "Bajos del Norte-10", "Alacranes-15", "Bajos del Norte-15", "Alacranes-25", "Bajos del Norte-25","Alacranes-35", "Bajos del Norte-35")

# reads in fst matrix
snpFstMa <- as.matrix(mx.fst$Fsts)
upperTriangle(snpFstMa, byrow=TRUE) <- lowerTriangle(snpFstMa)
snpFstMa <- snpFstMa[,pop.order] %>%
  .[pop.order,]
snpFstMa[upper.tri(snpFstMa)] <- NA
snpFstMa <- as.data.frame(snpFstMa)

snpFstMa$Pop = factor(row.names(snpFstMa), levels = unique(pop.order))

snpQMa <- as.matrix(mx.fst$Pvalues)
upperTriangle(snpQMa, byrow=TRUE) <- lowerTriangle(snpQMa)
snpQMa <- snpQMa[,pop.order] %>%
  .[pop.order,]
snpQMa[upper.tri(snpQMa)] <- NA
snpQMa <- as.data.frame(snpQMa)
snpQMa$Pop = factor(row.names(snpQMa), levels = unique(pop.order))

snpFstMa$Pop = factor(row.names(snpFstMa), levels = unique(pop.order))
snpFst = melt(snpFstMa, id.vars = "Pop", value.name = "Fst", variable.name = "Pop2", na.rm = TRUE)
snpFst = snpFst[snpFst$Pop != snpFst$Pop2,]
snpFst$Fst = round(snpFst$Fst, 3)
snpFst = snpFst %>% mutate(Fst = replace(Fst, Fst < 0, 0))
head(snpFst)

snpQ = melt(snpQMa, id.vars = "Pop", value.name = "Pval", variable.name = "Pop2", na.rm = TRUE)
snpQ = snpQ[snpQ$Pop != snpQ$Pop2,]
snpQ$Qval = p.adjust(snpQ$Pval, method = "BH")
head(snpQ)

snpHeatmapA = ggplot(data = snpFst, aes(Pop, Pop2, fill = Fst))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "white", high = "red", midpoint = 0, limit = c(0, 0.08),
                       space = "Lab", name = expression(paste(italic("F")[ST])))+
  geom_text(data = snpFst, aes(Pop, Pop2, label = Fst), color = ifelse(snpQ$Qval <= 0.05,"black", "darkgrey"), size = ifelse(snpQ$Qval < 0.05, 6, 5), fontface = ifelse(snpQ$Qval <= 0.05, "bold", "plain")) +
  guides(fill=guide_colorbar(barwidth = 1, barheight = 12, title.position = "top", title.hjust = 0.5))+                     
  scale_y_discrete(position = "right", labels = c("Alacranes-10 m", "Bajos del Norte-10 m","Alacranes-15 m", "Bajos del Norte-15 m", "Alacranes-25 m", "Bajos del Norte-25 m","Alacranes-35 m"))+
  scale_x_discrete(labels = str_wrap(c("Bajos del Norte-10 m", "Alacranes-15 m", "Bajos del Norte-15 m", "Alacranes-25 m", "Bajos del Norte-25 m","Alacranes-35 m", "Bajos del Norte-35 m"), width = 4)) +
  #ggtitle("   SNP") +
  theme_minimal()

snpHeatmap = snpHeatmapA + theme(
  axis.text.x = element_text(vjust = 1, size = 16, hjust = 0.5, color = "black"),
  axis.text.y = element_text(size = 16, color = "black"),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.position = "right",
  legend.direction = "vertical",
  legend.title = element_text(size = 16),
  legend.text = element_text(size = 14),
  plot.title = element_text(size = 16)
)

snpHeatmap

ggsave("./figures/fig3.tiff", plot = snpHeatmap, width = 30, height = 15, units = "cm", dpi = 300)
```

## NGSAdmix Plot
```{r, ADMIXTURE}
paletteer_d("NineteenEightyR::miami2")
#colPal2=c("#FFFFCC", "#41B6C4")
colPal2=c("#3FB8AFFF", "#DAD8A7FF")
mxNgsAdmix <- read.csv("mxMcavNoClones_k2k3.csv") %>% arrange(site, depthZoneM, -cluster2.2)
# mxK2$sample = factor(mxK2$sample, levels = mxK2$sample[order(-mxK2$cluster2)])
sampleCounts = plyr::count(mxNgsAdmix, c('site','depthZoneM'))
meltedList = reshape2::melt(lapply(sampleCounts$freq,function(x){c(1:x)}))

mxNgsAdmix$barPlotOrder = meltedList$value
mxNgsAdmix$depthZoneM=as.factor(mxNgsAdmix$depthZoneM)
mdat = melt(mxNgsAdmix, id.vars=c("sample", "site", "depthZoneM", "barPlotOrder"), variable.name="Ancestry", value.name="Fraction")
#mdat2$site=as.factor(mdat2$site)
mdat$depthZoneM=as.factor(mdat$depthZoneM)
levels(mdat$depthZoneM)
levels(mdat$depthZoneM) = c("10 m", "15 m", "25 m", "35 m")

p2 = ggplot(data = subset(mdat, subset = mdat$Ancestry %in% c("cluster2.1", "cluster2.2")), aes(x=barPlotOrder, y=Fraction, fill=Ancestry, order=barPlotOrder)) +
  geom_bar(stat="identity", position="fill", width=1, colour="grey25") +
  #facet_grid(.~site, scales = "free", switch = "x", space = "free") +
  facet_grid(depthZoneM ~ site, scales = "free") + #faceting plots by Depth and Site
  labs(x = "Population", y = "Ancestry") +
  ggtitle(expression(paste(italic("K"), "=2"))) +
  theme(plot.title = element_text(size=40),
        panel.grid=element_blank(),
        panel.background=element_rect(fill=NA, colour="grey25"),
        panel.spacing.x=grid:::unit(0, "lines"),
        panel.border = element_rect(fill=NA,color="black", size=2, linetype="solid"),
        axis.text.x=element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        strip.background = element_rect(fill = "white", size = 0.9, color="black"),
        strip.text.x = element_text(size = 30),
        strip.text.y = element_blank(),
        #strip.text=element_text(size=20, angle=90),
        legend.key=element_blank(),
        legend.position = "none",
        legend.title = element_blank()) +
  scale_x_discrete(expand=c(0, 0)) +
  scale_y_continuous(expand=c(0, 0)) +
  #scale_fill_manual(values = brewer.pal(name = "YlGnBu", n = 4), name = "Cluster") +
  scale_fill_manual(values = colPal2, name = "Cluster") +
  guides(fill=guide_legend(override.aes=list(colour=NULL)))
p2

###########
#colPal3=c("#FFFFCC", "#41B6C4", "#225EA8")
colPal3=c("#3FB8AFFF", "#DAD8A7FF", "#FF3D7FFF")
# mxK3 <- read.csv("mxMcavNoClones_K3.csv") %>% arrange(pop, depthZoneM, -cluster2)
# # mxK3$sample = factor(mxK3$sample, levels = mxK3$sample[order(-mxK3$cluster2)])
# sampleCountsK3 = plyr::count(mxK3, c('pop','depthZoneM'))
# meltedListK3 = reshape2::melt(lapply(sampleCountsK3$freq,function(x){c(1:x)}))
# 
# mxK3$barPlotOrderK3 = meltedListK3$value
# mdat3 = melt(mxK3, id.vars=c("sample", "pop", "depthZoneM", "barPlotOrderK3"), variable.name="Ancestry", value.name="Fraction")
# #mdat2$pop=as.factor(mdat3$pop)
# mdat3$depthZoneM=as.factor(mdat3$depthZoneM)
# levels(mdat3$depthZoneM)
# levels(mdat3$depthZoneM) = c("10 m", "15 m", "25 m", "35 m")


p3 = ggplot(data = subset(mdat, subset = mdat$Ancestry %in% c("cluster3.1", "cluster3.2", "cluster3.3")), aes(x=barPlotOrder, y=Fraction, fill=Ancestry, order=barPlotOrder)) +
  geom_bar(stat="identity", position="fill", width=1, colour="grey25") +
  #facet_grid(.~pop, scales = "free", switch = "x", space = "free") +
  facet_grid(depthZoneM ~ site, scales = "free") + #faceting plots by Depth and Site
  labs(x = "Population", y = "Ancestry") +
  ggtitle(expression(paste(italic("K"), "=3"))) +
  theme(plot.title = element_text(size=40),
        panel.grid=element_blank(),
        panel.background=element_rect(fill=NA, colour="grey25"),
        panel.spacing.x=grid:::unit(0, "lines"),
        panel.border = element_rect(fill=NA,color="black", size=2, linetype="solid"),
        axis.text.x=element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        strip.background = element_rect(fill = "white", size = 0.9, color="black"),
        strip.text.x = element_text(size = 30),
        strip.text.y = element_text(size = 30),
        #strip.text=element_text(size=20, angle=90),
        legend.key=element_blank(),
        legend.position = "none",
        legend.title = element_blank()) +
  scale_x_discrete(expand=c(0, 0)) +
  scale_y_continuous(expand=c(0, 0)) +
  #scale_fill_manual(values = brewer.pal(name = "YlGnBu", n = 4), name = "Cluster") +
  scale_fill_manual(values = colPal3, name = "Cluster") +
  guides(fill=guide_legend(override.aes=list(colour=NULL)))
p3

combinedAdmix = (p2 | p3)
combinedAdmix

ggsave("./figures/fig4.tiff", plot = combinedAdmix, width = 70, height = 40, units = "cm", dpi = 300)
```

##ITS2-type Profiles
```{r, its2BarPlot}
its2Profs = read.csv("mxIts2ProfCounts.csv", header = TRUE, check.names = FALSE)
head(its2Profs)

its2Profs$depthZoneM = factor(its2Profs$depthZoneM, levels = c("10", "15", "25", "35"))
levels(its2Profs$depthZoneM)

its2Profs$site = factor(its2Profs$site, levels(its2Profs$site)[c(1,2)])
its2Profs = its2Profs[order(its2Profs$site, its2Profs$depthZoneM), ]
head(its2Profs)

its2Profs = its2Profs %>% arrange(site, depthZoneM,
                                  `G3b`,
                                  `G3am`,
                                  `C3/C3b`,
                                  `C3-C3fc-C40g-C3b-C3an-C21-C3fd-C3s-C3bb`,
                                  `C3an/C3-C21-C3bb`, 
                                  `C3-C21-C3gu-C3gy-C3gw-C3gx-C3gv-C3b`, 
                                  `C3-C3an-C3de-C21-C3bb-C3fc-C3b`,
                                  `C3-C3de-C3bb-C21ae-C21-C3an-C3s`, 
                                  `C3-C3fc-C21-C3b-C3an-C3fd-C3bb`, 
                                  `C3-C3de-C21-C3bb-C3an-C21ae-C65b-C3s`, 
                                  `C3-C21-C3an-C3fc-C3gz-C3b-C3bb`, 
                                  `C3-C21-C3fc-C3an-C65b-C3b-C3bb`)  
                      
                      

sampleCounts = plyr::count(its2Profs, c('site','depthZoneM'))
meltedList = reshape2::melt(lapply(sampleCounts$freq,function(x){c(1:x)}))
its2Profs$barPlotOrder = meltedList$value
its2Profs=its2Profs[c(1,ncol(its2Profs),2:(ncol(its2Profs)-1))]

colOrder2 = order(colSums(its2Profs[10:length(its2Profs[1,])]), decreasing = TRUE)+9 #Add the number of columns of metadata

its2ProfsPerc = cbind(its2Profs[,c(1:9)],its2Profs[,c(colOrder2)])
its2ProfsPerc$sum = apply(its2ProfsPerc[, c(10:length(its2ProfsPerc[1,]))], 1, function(x) {
  sum(x, na.rm = T)
})

its2ProfsPerc = cbind(its2ProfsPerc[, c(1:9)], (its2ProfsPerc[, c(10:(ncol(its2ProfsPerc)-1))] 
                                                        / its2ProfsPerc$sum))
head(its2ProfsPerc)

apply(its2ProfsPerc[, c(10:(ncol(its2ProfsPerc)))], 1, function(x) {
  sum(x, na.rm = T)
})

head(its2ProfsPerc)


gssProf = otuStack(its2ProfsPerc, count.columns = c(10:length(its2ProfsPerc[1, ])),
                       condition.columns = c(1:9)) # remove summ rows
gssProf = gssProf %>% filter(otu != "summ") %>% droplevels()

levels(gssProf$otu)

levels(gssProf$depthZoneM)
levels(gssProf$depthZoneM) = c("10 m", "15 m", "25 m", "35 m")
levels(gssProf$site)
levels(gssProf$site) = c("Alacranes", "Bajos del Norte")
levels(gssProf$depthZoneM)
levels(gssProf$site)

#gssProf %>% arrange(site, depthZoneM, count, otu)

#Construct bar plot

its2ProfsPlotA = ggplot(gssProf, aes(x = barPlotOrder, y = count, fill = factor(otu))) +
  geom_bar(position = "stack", stat = "identity", color = "black", size = 0.25) + 
  ylab("Proportion") +
  scale_fill_paletteer_d("dichromat::Categorical.12")+
  labs(fill = expression(paste(italic("ITS2"), " type profile"))) +
  guides(fill = guide_legend(ncol = 2, reverse = FALSE)) +
  facet_grid(depthZoneM ~ site, scales = "free_x") + #faceting plots by Depth and Site
  theme_bw()

its2ProfsPlot = its2ProfsPlotA +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "bottom",
        legend.title = element_text(color = "black", size = 12, hjust = 0.5, angle = 90),
        legend.text = element_text(color = "black", size = 10),
        legend.key = element_blank(),
        legend.key.size = unit(0.75,"line"),
        legend.background = element_blank(),
        panel.border = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.grid=element_blank(),
        plot.background = element_blank(),
        strip.text.x = element_text(size = 12),
        strip.text.y = element_text(size = 12),
        strip.background = element_rect(fill = "white", size = 0.9)
  )

its2ProfsPlot

ggsave("./figures/fig5.tiff", plot = its2ProfsPlot, width = 8.5, height = 8.5, unit = "in", dpi = 600)
```

##ITS2 Stats Analysis
```{r, itsStats}
set.seed(694) #setting seed allows repetition of randomized processes

its2dispS = betadisper(vegdist(its2Profs[, c(10:ncol(its2Profs))]), its2Profs$site)

anova(its2dispS)
#No significant effect of site on beta diversity

set.seed(694)

its2dispD = betadisper(vegdist(its2Profs[, c(10:ncol(its2Profs))]), its2Profs$depthZoneM)

anova(its2dispD)
#No significant effect of depth on beta diversity

#PERMANOVA
set.seed(694)
its2Adonis = adonis(its2Profs[, c(10:ncol(its2Profs))] ~ depthZoneM * site, 
                    data = its2Profs, permutations = 9999, method = "bray")

its2Adonis

set.seed(694)

its2PWAdonis = pairwise.adonis(its2Profs[, c(10:ncol(its2Profs))],
                               factors = its2Profs$depthZoneM,
                               sim.method = "bray", p.adjust.m = "BH", perm = 9999)

its2PWAdonis 
```

##ITS2 Symbiont PCoA with Bray Curtis
```{r, itsPCoA}
zooxProfs = its2Profs[,10:ncol(its2Profs)]
zooxDist=vegdist(zooxProfs, method="bray")

zooxMds = cmdscale(zooxDist, eig = TRUE, x.ret = TRUE)

# Determine percent variation captured on each axis
# Calculate the eigenvalues so later we can figure out % variation shown on each Principal Coordinate
mxZooxPcoaVar = round(zooxMds$eig/sum(zooxMds$eig)*100, 1)
mxZooxPcoaVar

# Format data to plot
mxZooxPcoaValues = zooxMds$points
mxZooxPcoaValues

zooxI2P = its2Profs[,1:9]
row.names(zooxI2P) = zooxI2P[,1]
mxZooxPcoaValues=cbind(zooxI2P, mxZooxPcoaValues)
mxZooxPcoaValues =as.data.frame(mxZooxPcoaValues, sample = rownames(mxZooxPcoaValues))
colnames(mxZooxPcoaValues)[10] <- "PCo1"
colnames(mxZooxPcoaValues)[11] <- "PCo2"
mxZooxPcoaValues$popDepth<- paste(mxZooxPcoaValues$site, "-", mxZooxPcoaValues$depthZoneM)
head(mxZooxPcoaValues)

zooxPCoA = merge(mxZooxPcoaValues, aggregate(cbind(mean.x=PCo1,mean.y=PCo2)~popDepth, mxZooxPcoaValues, mean), by="popDepth")

#snpPCoA$depthZoneM=as.factor(snpPCoA$depthZoneM)
#snpPCoA$depthZoneM = factor(snpPCoA$depthZoneM, levels(snpPCoA$depthZoneM)[c(2,1)])

# SNP PCoA biplot
mxZooxPcoaPlotA = ggplot(zooxPCoA, aes(x = PCo1, y = PCo2, color = site, fill = site, shape = depthZoneM, linetype = depthZoneM)) +
  geom_hline(yintercept = 0, color = "gray90", size = 0.5) +
  geom_vline(xintercept = 0, color = "gray90", size = 0.5) +
  #stat_ellipse(data = subset(zooxPCoA, type = "n", geom = "polygon")) + #ellipse
  scale_linetype_manual(values=c(1,2,4,3), name = "Depth Zone")+
  geom_point(aes(x = PCo1, y = PCo2, shape = depthZoneM), size = 3, alpha = 0.3, show.legend = FALSE, guide=FALSE, position=position_jitter()) + #individual's points indicated by circles
  #geom_jitter(alpha = 0.3, show.legend=FALSE)+
  scale_shape_manual(values = c(24,22,23,25), breaks=c("10","15", "25", "35"), name = "Depth Zone", labels=c("10 m","15 m", "25 m", "35 m")) +
  geom_point(aes(x = mean.x, y = mean.y, shape = depthZoneM), size = 5, color = "black") + #population centroids indicated by triangles
  #scale_fill_paletteer_d("LaCroixColoR::PassionFruit", breaks=c("Alacranes", "Bajos del Norte"), name= "Population")+
  scale_fill_manual(values = c("seagreen3", "blue"), name= "Population")+
  #scale_color_paletteer_d("LaCroixColoR::PassionFruit", breaks=c("Alacranes", "Bajos del Norte"), name= "Population", guide=FALSE) +
  scale_color_manual(values = c("seagreen3", "blue"), name= "Population")+
  xlab(paste ("PCo 1 (", mxZooxPcoaVar[1],"%)", sep = "")) + #Prints percent variation explained by first axis
  ylab(paste ("PCo 2 (", mxZooxPcoaVar[2],"%)", sep = "")) + #Prints percent variation explained by second axis
  guides(shape = guide_legend(order = 2), fill = guide_legend(override.aes = list(shape = 22, size = 4, color = NA), order = 1))+ #linetype = guide_legend(order = 3),
  theme_bw()

mxZooxPcoaPlot = mxZooxPcoaPlotA +
  theme(axis.title.x = element_text(color = "black", size = 10),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.title.y = element_text(color = "black", size = 10),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        legend.position = "left",
        panel.border = element_rect(color = "black", size = 1.2),
        panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())



mxZooxPcoaPlot

ggsave("./figures/fig6.tiff", plot = mxZooxPcoaPlot, height = 5, width = 7, units = "in", dpi = 300)
```